Using SP, igraph and networkplanner libraries to analyze a point dataset
------------------------------------------------------------------------

Load point dataset from csv and map it 
```{r load, fig.height=3, message=F, warning=F}
require(sp)
require(methods)
require(RColorBrewer)
csv_file <- "data/mmr_hh_spatial.csv"
hh_df <- read.csv(csv_file)

mapping_from <- c("household._gps_longitude", "household._gps_latitude", "Household_geocharacterization.hh_size", "Household_geocharacterization.hh_rooms")
mapping_to <- c("X", "Y", "hh_size", "hh_rooms")
hh_df[,mapping_to] <- hh_df[,mapping_from]
hh_df <- hh_df[,mapping_to]
# transform the points and hh_size from factors to numerics (NOT VERY INTUITIVE)
unfactor_hh_size <- function(hh_size) {
    hh_size_num <- suppressWarnings(as.numeric(levels(hh_size)))
    hh_size_num[hh_size]
}
    
hh_df <- transform(hh_df, hh_size=unfactor_hh_size(hh_size), 
                          X=as.numeric(levels(X))[X], 
                          Y=as.numeric(levels(Y))[Y])
hh_sp_df <- hh_df
coordinates(hh_sp_df) <- ~X + Y

# and set the projection
proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84"
hh_sp_df@proj4string <- CRS(proj)

# simple plot of points (colored by hh_size)
hh_classes <- with(hh_df, seq(min(hh_size, na.rm=T), max(hh_size, na.rm=T), len=5)) 
color_indices <- findInterval(hh_df$hh_size, hh_classes)
colors <- brewer.pal(5, "BuGn")[color_indices]
colors[is.na(colors)] <- "#FFFFFF"

plot(hh_sp_df, pch=21, col="#000000", bg=colors, lwd=0.2)
```

Calculate the MST via igraph and plot it 
```{r create_mst, fig.width=12, fig.height=12, message=F, warning=F}
require(igraph)
require(geosphere)

# create the full graph
edge_df <- data.frame(t(combn(1:nrow(hh_sp_df), 2)))
names(edge_df) <- c("from", "to")

# assign distances and create graph
edge_df$weight <- distHaversine(hh_sp_df[edge_df$from,], hh_sp_df[edge_df$to,])
g <- graph.data.frame(edge_df, directed=F)

# compute the MST
mst_g <- minimum.spanning.tree(g)
plot(mst_g, vertex.size=4, vertex.label="")

```

Calculate costs associated with creating an LV network
```{r calc_costs, message=F, warning=F}
network_span <- sum(E(mst_g)$weight)

# sum of distances between edges / num edges
mean_edge_len <- network_span/length(E(mst_g))

# assign costs to grid components
LV_POLE_DIST_M <- 50
LV_POLE_COST_D <- 150
LV_COST_PER_M_D <- 15
MV_TO_LV_XFRM_COST_D <- 10000
LV_SURVEY_COST_PER_M_D <- 0.0025

# edge cost function
edge_cost <- function(distance) {
    pole_cost <- (distance / LV_POLE_DIST_M) * LV_POLE_COST_D
    lv_cost <- distance * LV_COST_PER_M_D
    survey_cost <- distance * LV_SURVEY_COST_PER_M_D
    pole_cost + lv_cost + survey_cost
}

# order edges in desc length order
mst_edge_df <- get.data.frame(mst_g, what="edges")
mst_edge_df <- mst_edge_df[order(-mst_edge_df$weight),]
mst_edge_df <- transform(mst_edge_df, lv_cost=edge_cost(weight))

# assign costs to edges
E(mst_g)$lv_cost <- edge_cost(E(mst_g)$weight)

num_transformers <- 1
mst_based_cost <- sum(mst_edge_df$lv_cost) + (num_transformers * MV_TO_LV_XFRM_COST_D)

# summarize
cat("mean edge length (meters): ", mean_edge_len, "\n")
cat("mst based cost (dollars): ", mst_based_cost, "\n")
```

Construct the NetworkPlan and map it
```{r networkplan, fig.width=12, fig.height=3, message=F, warning=F}
# map points to vertices in graph
require(networkplanner)
np <- create_networkplan_from_graph(mst_g, hh_df, proj)

# get spatial dataframes to map
sp_dfs <- as_spatial_dataframes(np)
edge_sp_df <- sp_dfs$edges

plot(edge_sp_df)
```

Reduce costs by substituting lower cost transformers for expensive edges
```{r cost_reduction, message=F, warning=F}
# find the 'expensive' edges and remove 'em in favor of transformers
expensive_edge_df <- mst_edge_df[mst_edge_df$lv_cost > MV_TO_LV_XFRM_COST_D,]
reduced_mst_g <- mst_g
reduced_mst_g[from=expensive_edge_df$from, to=expensive_edge_df$to] <- F
num_transformers <- num_transformers + nrow(expensive_edge_df)

reduced_network_span <- sum(E(reduced_mst_g)$weight)

# sum of distances between edges / num edges
reduced_mean_edge_len <- reduced_network_span/length(E(reduced_mst_g))
    
reduced_mst_based_cost <- sum(E(reduced_mst_g)$lv_cost) + (num_transformers * MV_TO_LV_XFRM_COST_D)

# summarize
cat("reduced mean edge length (meters): ", reduced_mean_edge_len, "\n")
cat("reduced mst based cost (dollars): ", reduced_mst_based_cost, "\n")

```

Construct the Reduced NetworkPlan and map it
```{r reduced_networkplan, fig.width=12, fig.height=3, message=F, warning=F}
# map points to vertices in graph
np <- create_networkplan_from_graph(reduced_mst_g, hh_df, proj)

# get spatial dataframes to map
sp_dfs <- as_spatial_dataframes(np)
edge_sp_df <- sp_dfs$edges

plot(edge_sp_df)

```

Decompose the reduced network and map the largest component
```{r decompose, fig.width=12, fig.height=12, message=F, warning=F}
components <- decompose.graph(reduced_mst_g)
component_sizes <- sapply(components, function(g) { length(E(g)) })

largest_component <- components[[which(component_sizes==max(component_sizes))]]
# map points to vertices in graph
np <- create_networkplan_from_graph(largest_component, hh_df, proj)

# get spatial dataframes to map
sp_dfs <- as_spatial_dataframes(np)
edge_sp_df <- sp_dfs$edges

dist_classes <- with(edge_sp_df@data, seq(min(lv_cost), max(lv_cost), len=5)) 
color_indices <- findInterval(edge_sp_df@data$lv_cost, dist_classes)
colors <- brewer.pal(5, "BuGn")[color_indices]

plot(edge_sp_df, col=colors)
```
